---
title: "Getting Started with pgsmetrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with pgsmetrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pgsmetrics)
```

## Example (binary trait)

First, let's simulate some data:

```{r simulate_data}
n_simul <- 1e4
df <- simulate_data(n = n_simul, n_pgs = 5)
```

Now, let's use the `pgsmetrics` function to fit models with different PGS and compute performance metrics.
In this example, we'll use 5 PGS with varying degree of measurement error (TODO see `PGS simulation` vignette).
We assume a binary variable (`y_bin`), adjust for `age`, and use 100 bootstrap replicates (TODO $>200$ or $>500$ recommended).

```{r basic_usage}
results <- pgsmetrics(
  df,
  pgs = paste0("pgs", 1:5),
  dep = "y_bin",
  covars = c("age"),
  n_boot = 100,
  n_cores = 1
)
```

The metrics are stored under `results$metrics`.
There are three different values for each PGS and each metric, one for a covariate-only model (no PGS, `cov`), one for a full model (`full`), and one for the difference between `full` and `cov` (incremental value, `full - cov`).

```{r print_results}
print(results, n = 1)
```

Let's plot the metrics

```{r plot_results, fig.height=2.5}
plot(results, model_type = "partial", ncol = 5)
```

In each bootstrap replication, we rank the PGS based on the metrics and compute confidence intervals.
Let's plot the ranks.

```{r plot_results_ranks, fig.height=2.5}
plot(results, type = "ranks", ncol = 5)
```

Diff

```{r plot_results_diff, fig.height=4}
plot(results, type = "diff", ncol = 5)
```

## Example (continuous trait)

```{r results_quant}
results_quant <- pgsmetrics(
  df,
  pgs = paste0("pgs", 1:5),
  dep = "y",
  covars = c("age"),
  n_boot = 100,
  n_cores = 1
)
```

The metrics are stored under `results_quant$metrics`

```{r print_results_quant}
print(results_quant, n = 1)
```

Let's plot the metrics

```{r plot_results_quant, fig.height=2.5}
plot(results_quant, ncol = 5)
```

Let's plot the ranks

```{r plot_results_quant_ranks, fig.height=2.5}
plot(results_quant, type = "ranks", ncol = 5)
```

Diff

```{r plot_results_quant_diff, fig.height=4}
plot(results_quant, type = "diff", ncol = 4)
```

## Calibration

TODO text.

```{r calibration_plots, fig.height=4}
p_cal1 <- plot_calibration(df, pgs = "pgs1", dep = "y_bin")
p_cal2 <- plot_calibration(df, pgs = "pgs1", dep = "y_bin", covars = c("age", "sex"))
cowplot::plot_grid(p_cal1, p_cal2, nrow = 1)
```

## Net benefit / decision curves

TODO text.

```{r dcurve_plots, fig.height=4}
p_dcurve1 <- plot_dcurve(df, pgs = "pgs1", dep = "y_bin")
p_dcurve2 <- plot_dcurve(df, pgs = "pgs1", dep = "y_bin", covars = c("age", "sex"))
cowplot::plot_grid(p_dcurve1, p_dcurve2, nrow = 1)
```
